19Q1

### 2019-04-02 19Q1.Rmd
update 19Q1.Rmd and knit to HTML

## CCLE genomic annotations

### Mutation status (Mut)

@@

### Copy number (CN)

@@

### Gene expression (GE)

@@

@@

# Account for cell line aliases and missing histologies
auc_unmatched_na_histologies <- read.delim("./data_munging/pozdeyev2016/auc_unmatched_na_histologies.csv", sep = ",", row.names = 1)
auc_unmatched_na_histologies$accession_id <- NULL
auc_unmatched_na_histologies$Notes <- NULL
auc_unmatched_na_histologies$Histology <- ifelse(auc_unmatched_na_histologies$Histology == "", NA, as.character(auc_unmatched_na_histologies$Histology))
auc_unmatched_na_histologies$CCLE_ccl_name <- ifelse(auc_unmatched_na_histologies$CCLE_ccl_name  == "", NA, as.character(auc_unmatched_na_histologies$CCLE_ccl_name ))
auc_unmatched_na_histologies$Broad_ID <- ifelse(auc_unmatched_na_histologies$Broad_ID == "", NA, as.character(auc_unmatched_na_histologies$Broad_ID))

# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_unmatched_na_histologies)
```

@@

## Matched data

```{r, eval=TRUE}
auc_matched <- read.delim("./data_munging/pharmacogenomics_raw_files/celline_stack.csv.gz", sep = ",")
auc_matched <- merge(ccl_histologies, auc_matched , by = "Cell_line", all.y = TRUE)
num_auc_per_db_matched <- auc_matched %>% count(Database)
ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line))
num_ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Database)
# only CCLs represetented in 2 or 3 databases were included
num_db_per_ccl_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Cell_line)
```

There were `r length(setdiff(unique(ccl_histologies$Cell_line), unique(auc_matched $Cell_line)))` cell lines not represented in the Pozdeyev et al. harmonized dataset.

@@

## Unmatched data

```{r, eval=FALSE}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
ccle_poz <- merge(ccl_histologies, ccle_poz, by = "Cell_line", all.y = TRUE)
ccle_poz$InCCLE <- "1"
ccle_poz$InCTRP <- ""
ccle_poz$InGDSC <- ""

ctrp_poz <- read.delim("./../htscreens_giant_files/pozdeyev2016/CTRP_drug_response_metrics.csv.gz", sep = ",")
ctrp_poz <- merge(ccl_histologies, ctrp_poz, by = "Cell_line", all.y = TRUE)
ctrp_poz$InCCLE <- ""
ctrp_poz$InCTRP <- "1"
ctrp_poz$InGDSC <- ""
# ctrp_test <- merge(select(ctrp_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CTRP"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

gdsc_poz <- read.delim("./data_munging/pozdeyev2016/GDSC_drug_response_metrics.csv.gz", sep = ",")
gdsc_poz <- merge(ccl_histologies, gdsc_poz, by = "Cell_line", all.y = TRUE)
gdsc_poz$InCCLE <- ""
gdsc_poz$InCTRP <- ""
gdsc_poz$InGDSC <- "1"
# gdsc_test <- merge(select(gdsc_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "GDSC"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

auc_unmatched <- bind_rows(ccle_poz, ctrp_poz, gdsc_poz)
auc_unmatched$WasDrugScreened <- 1
# auc_unmatched$AUC_EC50 <- as.numeric(as.character(auc_unmatched$AUC_EC50))
auc_unmatched$AUC_IC50 <- as.numeric(as.character(auc_unmatched$AUC_IC50))
# auc_unmatched$sample_ID <- with(auc_unmatched, paste(Database, Cell_line, Histology, sep = "_"))
auc_unmatched$Drug_name_Database <- with(auc_unmatched, paste(Drug_name, Database, sep = "_"))

num_auc_per_db_unmatched <- auc_unmatched %>% count(Database)

# Identify cell lines without histology indications. Write to file to edit for merging back to ccl_histologies file
# test <- filter(auc_unmatched, is.na(Histology))[, 1:10]
# auc_unmatched_na_hists <- filter(auc_unmatched, is.na(Histology))$Cell_line
# write.csv(auc_unmatched_na_hists, "./data_munging/pozdeyev2016/auc_unmatched_na_histologies.csv")

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_unmatched_wide <- dcast(auc_unmatched, Cell_line + Histology + CCLE_ccl_name + Broad_ID + InCCLE + InCTRP + InGDSC + WasDrugScreened ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_unmatched_wide[auc_unmatched_wide == 1000000] <- ""
auc_unmatched_wide <- auc_unmatched_wide %>% group_by(Cell_line, Histology, CCLE_ccl_name, Broad_ID, WasDrugScreened) %>% summarise_all(funs(trimws(paste(., collapse = ""))))
auc_unmatched_wide$InCCLE[auc_unmatched_wide$InCCLE == ""] <- 0
auc_unmatched_wide$InCTRP[auc_unmatched_wide$InCTRP == ""] <- 0
auc_unmatched_wide$InGDSC[auc_unmatched_wide$InGDSC == ""] <- 0
auc_unmatched_wide[auc_unmatched_wide == ""] <- NA
auc_unmatched_wide <- add_column(auc_unmatched_wide, WasCRISPRScreened = NA, .after = "WasDrugScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasCopyNumberData = NA, .after = "WasCRISPRScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasGeneExpressionData = NA, .after = "WasCRISPRScreened")
auc_unmatched_wide <- add_column(auc_unmatched_wide, HasMutationData = NA, .after = "WasCRISPRScreened")

ccls_not_drugscreened <- filter(ccl_histologies, Cell_line %in% setdiff(ccl_histologies$Cell_line, auc_unmatched_wide$Cell_line))
ccls_not_drugscreened$WasDrugScreened <- 0

maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")

auc_all_ccls <- bind_rows(auc_unmatched_wide, ccls_not_drugscreened)
auc_all_ccls$HasMutationData <- with(auc_all_ccls, ifelse(Broad_ID %in% maf_df$Broad_ID, 1, 0))
auc_all_ccls$HasCopyNumberData <- with(auc_all_ccls, ifelse(Broad_ID %in% cn_melt$Broad_ID, 1, 0))
auc_all_ccls$HasGeneExpressionData <- with(auc_all_ccls, ifelse(Broad_ID %in% ge_melt$Broad_ID, 1, 0))
auc_all_ccls$WasCRISPRScreened <- with(auc_all_ccls, ifelse(Broad_ID %in% crispr_meta$Broad_ID, 1, 0))
auc_all_ccls$InCCLE <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCCLE))
auc_all_ccls$InCTRP <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCTRP))
auc_all_ccls$InGDSC <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InGDSC))
auc_all_ccls[, -c(1:4)] <- apply(auc_all_ccls[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all_ccls <- subset(auc_all_ccls, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all_ccls, file = "./data_munging/pozdeyev2016/merged_auc_v20190327.3.csv.gz", row.names = FALSE, na = "NA")
```

```{r}
# test <- as.data.frame(table(auc_all_ccls$Cell_line))
# test <- filter(auc_all_ccls, Cell_line %in% c("KMH2", "MS1", "TT"))[, 1:12]

length(unique(sapply(strsplit(as.character(colnames(auc_all_ccls)[13:length(colnames(auc_all_ccls))]), split = "_"), function(x) x[1])))
```

### README

Data source:
- Pozdeyev, N., Yoo, M., Mackie, R., Schweppe, R. E., Tan, A. C., & Haugen, B. R. (2016). Integrating heterogeneous drug sensitivity data from cancer pharmacogenomic studies. Oncotarget, 7(32), 51619–51625. https://doi.org/10.18632/oncotarget.10010
- Online visualization tool: http://tanlab.ucdenver.edu/QAPC/
- Zipped file containing their R scripts and full datasets can be downloaded here: http://tanlab.ucdenver.edu/QAPC/Downloads/Drug_Sensitvity_Metrics.zip

Column descriptions in merged_auc_v20190319.1.csv.gz:
- Cell_line: the name of the cancer cell line (CCL), harmonized across datasets by Pozdeyev et al. (2016)
- Histology: the general histology/tissue type of the CCL according to the CCLE annotations (see DepMap-2019q1-celllines_v2.csv file from the DepMap Data Portal)
- CCLE_ccl_name: the cell line name using CCLE nomenclature, which is a concatenation of the cell line name with the CCLE-annotated histology in all caps, separated by underscores
- Broad_ID: the custom identifiers (formatted as ACH-######) used by the Broad's Cancer Dependency Map project to ease cell line name harmonization efforts 
- WasDrugScreened: 1 means the cell was screened in at least one pharmacogenomic screens, and 0 means it was not screened in any of the pharmacogenomic screens
- WasCRISPRScreened: 1 means the cell line was screened in the Avana CRISPR screen, and 0 means it was not
- HasMutationData: 1 means the cell line had at least one recorded mutation in the MAF file (see file version below), and 0 means it was not annotated in the MAF file
- HasGeneExpressionData: 1 means gene expression data is available for the cell line (see file version below), and 0 means it does not have gene expression information
- HasCopyNumberData: 1 means copy number data is available for the cell line (see file version below), and 0 means it does not have copy number information
- InCCLE: 1 means the cell line was screened in CCLE, 0 means it was not
- InCTRP: 1 means the cell line was screened in CTRP, 0 means it was not
- InGDSC: 1 means the cell line was screened in GDSC, 0 means it was not
- All other columns: compound and drug names that were screened in at least one CCL in at least one database, formatted as Drug_DatabaseName
    - NA values indictate that compound wasn't screened in the respective CCL in the respective source dataset
    - Values are in AUC_IC50, the area under rhe dose response curve calculated from IC50 model

merged_auc_v20190319.1.csv.gz
merged_auc_v20190319.2.csv.gz:
- NAs in Histology column (excluding most of the cell lines indicated as being from a "collaborator" in CTRP) hand-curated by searches in DepMap, COSMIC, Cellosaurus, ATCC, canSAR Black, and Cell Model Passports
merged_auc_v20190327.3.csv.gz:
- Reformatted to have only unique cell lines in rows
- Cancer cell lines screened in more than one database have multiple AUC columns (column names are formatted as Drug_DatabaseName)
- Added information on binary presence/absencee of omic annotations for each cell line (i.e. mutations, copy number, gene expression)
- Added binary columns indicating presence in CCLE, CTRP, GDSC, drug screens in general (in 1, 2, or 3), and the Avana CRISPR screen
- Added cell lines that have at least one type of omic data but are not in the CRISPR or pharmacogenomic screens

Exploratory plots:
```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_all_ccl_unmatched_pozdeyev <- venn.diagram(list("CCLE" = unique(ccle_poz$Cell_line), "CTRP" = unique(ctrp_poz$Cell_line), "GDSC" = unique(gdsc_poz$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
# png("./gcn_denise/pozdeyev_unmatched/pozdeyev_venn_all_ccl_unmatched.png", res = 500, units = "in", width = 6, height = 6)
# grid.newpage()
# grid.draw(venn_all_ccl_unmatched_pozdeyev)
# dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_all_drugs_unmatched_pozdeyev <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
# png("./gcn_denise/pozdeyev_unmatched/pozdeyev_venn_all_drugs_unmatched.png", res = 500, units = "in", width = 6, height = 6)
# grid.newpage()
# grid.draw(venn_all_drugs_unmatched_pozdeyev)
# dev.off()

num_ccl_per_hist_per_db_unmatched_bar <- ggplot(data = unique(select(auc_unmatched, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
num_ccl_per_hist_per_db_unmatched_bar
# ggsave("./gcn_denise/pozdeyev_unmatched/pozdeyev_ccl_per_histology_per_db_unmatched.png", num_ccl_per_hist_per_db_unmatched_bar, device = "png", dpi = 450, width = 12, height = 8, units = "in")

>>>>

## Omics

@@

@@

@@

Load Rds:
```{r}
maf_df <- readRDS("./data_munging/19Q1/maf_df_19Q1.rds")
cn_melt <- readRDS("./data_munging/19Q1/cn_melt_19Q1.rds")
ge_melt <- readRDS("./data_munging/19Q1/ge_melt_19Q1.rds")
```

@@

# Account for cell line aliases and missing histologies
auc_na_histologies <- read.delim("./data_munging/pozdeyev2016/auc_na_histologies.csv", sep = ",", row.names = 1)

auc_na_histologies$accession_id <- NULL
auc_na_histologies$Notes <- NULL
auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_ccl_name <- ifelse(auc_na_histologies$CCLE_ccl_name  == "", NA, as.character(auc_na_histologies$CCLE_ccl_name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
```

@@

## Data wrangling

```{r, eval=FALSE}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")

ctrp_poz <- read.delim("./../htscreens_giant_files/pozdeyev2016/CTRP_drug_response_metrics.csv.gz", sep = ",")
# ctrp_test <- merge(select(ctrp_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CTRP"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

gdsc_poz <- read.delim("./data_munging/pozdeyev2016/GDSC_drug_response_metrics.csv.gz", sep = ",")
# gdsc_test <- merge(select(gdsc_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "GDSC"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

auc <- bind_rows(ccle_poz, ctrp_poz, gdsc_poz)

# Harmonize Pozdeyev cell line names with DepMap
auc$Cell_line <- toupper(gsub("-", "", auc$Cell_line))
auc$Cell_line <- ifelse(auc$Cell_line == "K052", "KO52", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "OVCAR3", "NIHOVCAR3", auc$Cell_line)
auc$Cell_line <- ifelse(auc$Cell_line == "TI73", "T173", auc$Cell_line)
auc <- merge(ccl_histologies, auc, by = "Cell_line", all.y = TRUE)

auc$WasDrugScreened <- 1
ccls_not_drugscreened <- filter(ccl_histologies, Cell_line %in% setdiff(ccl_histologies$Cell_line, auc_wide$Cell_line))
ccls_not_drugscreened$WasDrugScreened <- 0

auc$AUC_IC50 <- as.numeric(as.character(auc$AUC_IC50))
# auc$sample_ID <- with(auc, paste(Database, Cell_line, Histology, sep = "_"))
auc$Drug_name_Database <- with(auc, paste(Drug_name, Database, sep = "_"))

num_auc_per_db <- auc %>% count(Database)

# Identify cell lines without histology indications. Write to file to edit for merging back to ccl_histologies file
# test <- filter(auc, is.na(Histology))[, 1:10]
# auc_na_hists <- unique(filter(auc, is.na(Histology))$Cell_linet)
# write.csv(auc_na_hists, "./data_munging/pozdeyev2016/auc_na_histologies.csv")

# Create merged dataset with sample_ids in the rows and drugs across the columns
auc_wide <- dcast(auc, Cell_line + Histology + CCLE_ccl_name + Broad_ID + WasDrugScreened ~ Drug_name_Database, value.var = "AUC_IC50", fun.aggregate = mean, fill = 1000000)
auc_wide[auc_wide == 1000000] <- ""
auc_wide <- auc_wide %>% group_by(Cell_line, Histology, CCLE_ccl_name, Broad_ID, WasDrugScreened) %>% summarise_all(funs(trimws(paste(., collapse = ""))))
auc_wide[auc_wide == ""] <- NA

auc_all <- bind_rows(auc_wide, ccls_not_drugscreened)

auc_all <- add_column(auc_all, InCCLE = ifelse(auc_all$Cell_line %in% unique(ccle_poz$Cell_line), 1, 0), .after = "Broad_ID")
auc_all <- add_column(auc_all, InCTRP = ifelse(auc_all$Cell_line %in% unique(ctrp_poz$Cell_line), 1, 0), .after = "Broad_ID")
auc_all <- add_column(auc_all, InGDSC = ifelse(auc_all$Cell_line %in% unique(gdsc_poz$Cell_line), 1, 0), .after = "Broad_ID")

auc_all <- add_column(auc_all, WasCRISPRScreened = ifelse(auc_all$Broad_ID %in% crispr_meta$Broad_ID, 1, 0), .after = "WasDrugScreened")
auc_all <- add_column(auc_all, HasCopyNumberData = ifelse(auc_all$Broad_ID %in% unique(cn_melt$Broad_ID)[!is.na(unique(cn_melt$Broad_ID))], 1, 0), .after = "WasCRISPRScreened")
auc_all <- add_column(auc_all, HasGeneExpressionData = ifelse(auc_all$Broad_ID %in% unique(ge_melt$Broad_ID)[!is.na(unique(ge_melt$Broad_ID))], 1, 0), .after = "WasCRISPRScreened")
auc_all <- add_column(auc_all, HasMutationData = ifelse(auc_all$Broad_ID %in% unique(maf_df$Broad_ID)[!is.na(unique(maf_df$Broad_ID))], 1, 0), .after = "WasCRISPRScreened")

# auc_all[, -c(1:4)] <- apply(auc_all[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all <- subset(auc_all, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all, file = "./data_munging/pozdeyev2016/merged_auc_v20190402.4.csv.gz", row.names = FALSE, na = "NA")

# test <- as.data.frame(table(auc_all$Cell_line))
# test <- filter(auc_all, Cell_line %in% c("KMH2", "MS1", "TT"))[, 1:12]
```

```{r}
auc_all <- read.delim("./data_munging/pozdeyev2016/merged_auc_v20190402.4.csv.gz", sep = ",", check.names = FALSE)

# Unique drugs across all datasets
length(unique(sapply(strsplit(as.character(colnames(auc_all)[13:length(colnames(auc_all))]), split = "_"), function(x) x[1])))

sum(auc_all$HasCopyNumberData)
sum(auc_all$HasMutationData)
sum(auc_all$HasGeneExpressionData)
sum(auc_all$WasDrugScreened)

test <- filter(auc_all, HasMutationData == 1)[, 1:12]

pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)

ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)

all_omics_ccls <- intersect(intersect(unique(ge_ccls$Cell_line), unique(cn_ccls$Cell_line)), unique(mut_ccls$Cell_line))
sum(filter(auc_all, Cell_line %in% all_omics_ccls)$WasDrugScreened)

ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

auc_all %>% count(Histology)
```

## Plots

```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_ccl_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_ccl_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 6, units = "in")

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_drugs_overlap <- venn.diagram(list("CCLE" = unique(ccle_poz$Drug_name), "CTRP" = unique(ctrp_poz$Drug_name), "GDSC" = unique(gdsc_poz$Drug_name)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_drugs_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_drugs_overlap)
dev.off()

bar_ccl_per_hist_per_db <- ggplot(data = unique(select(auc, Database, Cell_line, Histology))) +
  facet_wrap(~ Database) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  # coord_cartesian(ylim = c(0, 130)) +
  scale_y_continuous(breaks = seq(0, 160, by = 20), labels = seq(0, 160, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology, grouped by database") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist_per_db
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology_per_db.png", bar_ccl_per_hist_per_db, device = "png", dpi = 450, width = 12, height = 8, units = "in")

bar_ccl_per_hist <- ggplot(data = unique(select(auc_all, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 220, by = 20), labels = seq(0, 220, by = 20)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all annotated cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_ccl_per_hist
ggsave("./gcn_denise/pozdeyev_unmatched/barplot_ccl_per_histology.png", bar_ccl_per_hist, device = "png", dpi = 450, width = 12, height = 8, units = "in")

histo_aucic50_per_db <- ggplot(data = auc) +
  facet_wrap(~ Database, scales = "free_y") +
  geom_histogram(aes(x = AUC_IC50, fill = Database), color = "black", alpha = 0.7, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_fill_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  guides(fill = "none") +
  labs(x = "AUC IC50", y = "Frequency", title = "Distributions of AUC IC50 values for all drugs in each dataset (unmatched)")
histo_aucic50_per_db
ggsave("./gcn_denise/pozdeyev_unmatched/histogram_aucic50_per_db.png", histo_aucic50_per_db, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

## README

DATA SOURCE

- Pozdeyev, N., Yoo, M., Mackie, R., Schweppe, R. E., Tan, A. C., & Haugen, B. R. (2016). Integrating heterogeneous drug sensitivity data from cancer pharmacogenomic studies. Oncotarget, 7(32), 51619–51625. https://doi.org/10.18632/oncotarget.10010
- Online visualization tool: http://tanlab.ucdenver.edu/QAPC/
- Zipped file containing their R scripts and full datasets can be downloaded here: http://tanlab.ucdenver.edu/QAPC/Downloads/Drug_Sensitvity_Metrics.zip

COLUMN DESCRIPTIONS

For latest version of merged_auc dataframe:

- Cell_line: the name of the cancer cell line (CCL), harmonized across datasets by Pozdeyev et al. (2016)
- Histology: the general histology/tissue type of the CCL according to the CCLE annotations (see DepMap-2019q1-celllines_v2.csv file from the DepMap Data Portal)
- CCLE_ccl_name: the cell line name using CCLE nomenclature, which is a concatenation of the cell line name with the CCLE-annotated histology in all caps, separated by underscores
- Broad_ID: the custom identifiers (formatted as ACH-######) used by the Broad's Cancer Dependency Map project to ease cell line name harmonization efforts 
- InCCLE: 1 means the cell line was screened in CCLE, 0 means it was not
- InCTRP: 1 means the cell line was screened in CTRP, 0 means it was not
- InGDSC: 1 means the cell line was screened in GDSC, 0 means it was not
- WasDrugScreened: 1 means the cell was screened in at least one pharmacogenomic screens, and 0 means it was not screened in any of the pharmacogenomic screens
- WasCRISPRScreened: 1 means the cell line was screened in the Avana CRISPR screen, and 0 means it was not
- HasMutationData: 1 means the cell line had at least one recorded mutation in the MAF file (see file version below), and 0 means it was not annotated in the MAF file
- HasGeneExpressionData: 1 means gene expression data is available for the cell line (see file version below), and 0 means it does not have gene expression information
- HasCopyNumberData: 1 means copy number data is available for the cell line (see file version below), and 0 means it does not have copy number information
- All other columns: compound and drug names that were screened in at least one CCL in at least one database, formatted as Drug_DatabaseName
    - NA values indictate that compound wasn't screened in the respective CCL in the respective source dataset
    - Values are in AUC_IC50, the area under rhe dose response curve calculated from IC50 model

CHANGE LOG FOR MERGED_AUC DATAFRAME

merged_auc_v20190319.1.csv.gz

merged_auc_v20190319.2.csv.gz:
- NAs in Histology column (excluding most of the cell lines indicated as being from a "collaborator" in CTRP) hand-curated by searches in DepMap, COSMIC, Cellosaurus, ATCC, canSAR Black, and Cell Model Passports

merged_auc_v20190327.3.csv.gz:
- Reformatted to have only unique cell lines in rows
- Cancer cell lines screened in more than one database have multiple AUC columns (column names are formatted as Drug_DatabaseName)
- Added information on binary presence/absencee of omic annotations for each cell line (i.e. mutations, copy number, gene expression)
- Added binary columns indicating presence in CCLE, CTRP, GDSC, drug screens in general (in 1, 2, or 3), and the Avana CRISPR screen
- Added cell lines that have at least one type of omic data but are not in the CRISPR or pharmacogenomic screens

merged_auc_v20190402.4.csv.gz:
- Fixed duplicate cell line rows derived from manually curating some of the CCL histologies that had been added in merged_auc_v20190319.2.csv.gz
- Fixed HasMutationData, HasCopyNumberData, and HasGeneExpression data columns
- Reordered columns

# Compare to TCGA

Quality check: cluster TCGA samples and cell lines—verify that same tumor types cluster together rather than clustering by cell lines and tumor samples (Kaiqiao)

Calculate Pearson correlations between data sets and create a heatmap (also look at Barretina et al. supplement for PCA and mutation load correlations)

```{r, eval=FALSE}
cn_seg <- read.delim("./data_munging/19Q1/copy_number_20190328.csv.gz", sep = ",")
tcga_pancan <- read.delim("~/Desktop/GDC-PANCAN.masked_cnv.tsv.gz", sep = "\t")

ggplot(cn_seg, aes(x = log2(Segment_Mean/2))) + 
  geom_density()

ggplot(tcga_pancan, aes(x = value)) + 
  geom_density()
```

# (OLD) Pozdeyev et al (2016) matched data

```{r, eval=FALSE}
auc_matched <- read.delim("./data_munging/pharmacogenomics_raw_files/celline_stack.csv.gz", sep = ",")
auc_matched <- merge(ccl_histologies, auc_matched , by = "Cell_line", all.y = TRUE)
num_auc_per_db_matched <- auc_matched %>% count(Database)
ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line))
num_ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Database)
# only CCLs represetented in 2 or 3 databases were included
num_db_per_ccl_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Cell_line)
```

@@

# (OLD) Denise's GCN dataframes
